\vspace{-.2in} Squint: Lossy Hierarchical Compression on Symbol Streams \vspace{-.1in}

### 2012-12-12 tex/paper.tex
Various edits up through 3.2

\documentclass[11pt]{article}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{fancyhdr,fancyvrb}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{natbib}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{xspace}

@@

@@

In summary, established work on grammar inference algorithms provides a solid
set of tools for experimenting with the kinds of local modifications we propose
to make to improve grammar structure by introducing lossiness.

@@

Work has also been done regarding the lossy compression of particular input
formats.  For text, methods such as replacing words with synonyms
\cite{semanticlossy} or reordering the middle letters of a word (i.e., all
letters save for the first and last) \cite{semilossless} have been explored.
However, these are very specific tricks to optimize text, and not more broadly
applicable to compressing hierarchical streams.

@@

In order to make our algorithm for lossily compressing streams as general as
possible, we decided to approach the problem of lossy compression by starting
with a lossless grammar inference algorithm.  By building off of a grammar
inference algorithm, we hope to take advantage of some of the partial
structures that are visible in a CFG representation of a stream.  In this
project, we chose \Sequitur, but any way of obtaining a grammar could be used.
Then, transformed the grammar, introducing loss and perform compressions by
simplifying the grammar. Thus our compression system is a multi-step process:

\begin{itemize}
  \item Infer a CFG representation of a stream (possibly using \Sequitur)
  \item Simplify the CFG, possibly introducing loss
  \item Output the CFG
\end{itemize}

Our core new contribution comes in step two, in what we shall call a
\emph{lossifier} algorithm, i.e. an algorithm that takes an admissible CFG and
simplifies it. The output CFG should be smaller than the input CFG but generate
a string which preserves the structure in the original string as much as
possible.

Building off the ideas in lossy compression discussed earlier, one idea for
simplifying data is to look at a part of the data and try to match it as
closely as possible with an existing set of primitives. In particular, just as
with Mpeg and fractal compression, to approximate one part of the grammar we
can use \emph{other parts of the grammar}. This is the seed for both of our
algorithms. Given a rule $S \rightarrow rhs$, we will try to replace occurences
of $S$ with other similar variables.

For example, consider the grammars in figure \ref{sred}.

@@

To simplify the grammar, it would make sense to replace B with A, since they
form two almost identical halves of S and are similar.

@@

To define whether two variables are \emph{similar}, we can look at their
expansions. Let $expand(rhs)$ denote the string of terminal symbols we get when
we recursively substitute all of the variables inside $rhs$. Then $V_1
\rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when

\[
  \frac{levenshtein(expand(rhs_1),expand(rhs_2))}
      {max(len(expand(rhs_1)),len(expand(rhs_2)))} < \epsilon
\]

where $levenshtein$ is the levenstein edit distance between two strings.  For
convenience we can define.

\[ lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))} \]

The first algorithm takes the idea of variable similarity and uses it to
simplify the grammar sequentially as much as possible.

>>>>

\documentclass[11pt]{article}
\usepackage[letterpaper,text={7in,9in},centering]{geometry}
\usepackage{tocloft}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array}
\usepackage{fancyhdr,fancyvrb}
\usepackage[pdftex]{graphicx}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{natbib}
\usepackage[lofdepth,lotdepth]{subfig}
\usepackage{sectsty}
\usepackage{setspace}
\usepackage{stackrel}
\usepackage{xspace}

@@

%\setlength{\cftbeforesecskip}{0pt}

@@

In summary, established work on grammar inference algorithms provides a solid
set of tools for experimenting with the kinds of local lossy modifications we
propose to make to simplify grammar structure.

@@

Work has also been done regarding the lossy compression of other input formats.
For text, methods such as replacing words with synonyms \cite{semanticlossy} or
reordering the middle letters of a word (i.e., all letters save for the first
and last) \cite{semilossless} have been explored.  However, these are very
specific tricks to optimize text, and not more broadly applicable to
compressing hierarchical streams.

@@

In order to make our algorithm for lossily compressing streams as general as
possible, we decided to approach the problem of lossy compression by starting
with a lossless grammar inference algorithm.  By building off of a grammar
inference algorithm, we hope to take advantage of some of the structures that
are visible in a CFG representation of a stream.  In this project, we chose
\Sequitur, but any method of inferring a grammar could be used.  After
inference, we simplified the resulting grammar, introducing loss and obtaining
greater compression.  Thus our full lossy grammar compression system is a
multi-step process:

\begin{itemize}
  \item Infer a CFG representation of a stream (we use \Sequitur).
  \item Simplify the CFG, introducing loss.
  \item Encode and output the CFG.
\end{itemize}

Our core contribution comes in step two, via what we refer to as a
\emph{lossifier} algorithm---an algorithm that takes an admissible CFG and
structurally simplifies it.  The output CFG should be smaller than the input
CFG but generate a string which preserves the structure in the original string
as much as possible.  In this paper, we focus on implementing and evaluating
this step, and hence omit step 3 altogether.

Building off the ideas in lossy compression discussed earlier, one idea for
simplifying data is to look at a part of the data and try to match it as
closely as possible with an existing set of primitives.  In particular, just as
with MPEG and fractal compression, to approximate one part of the grammar we
can use \emph{other parts of the grammar}.  This is the core principle behind
both of our algorithms.  Given a rule $A \rightarrow \alpha_1, \ldots,
\alpha_n$, we will try to replace occurences of $A$ with other similar
production rules.

For example, consider the first grammar given in \autoref{sred}.  In this
graphical representation, we omit the ordering of symbols on the right-hand
side of production rules.  The picture makes it clear, however, that one easy
route for simplifying the grammar is to simply replace the rule B with A.  They
differ in their expansions by only a single character, `e', and they form
almost identical halves of S.

@@

@@

To define whether two variables are \emph{similar}, we can look at their
expansions. Let $expand(rhs)$ denote the string of terminal symbols we get when
we recursively substitute all of the variables inside $rhs$. Then $V_1
\rightarrow rhs_1$ and $V_2 \rightarrow rhs_2$ are similar when
\[
  \frac{levenshtein(expand(rhs_1),expand(rhs_2))}
      {max(len(expand(rhs_1)),len(expand(rhs_2)))} < \epsilon
\]
where $levenshtein$ is the Levenstein edit distance between two strings.  For
convenience we can define.

\[ lsim(s1,s2) = \frac{levenshtein(s1,s2)}{max(len(s1),len(s2))} \]
Our first algorithm takes the idea of variable similarity and uses it to
iteratively simplify the grammar.
